---
title: "First assignment.- Data Visualization"
output: html_document
date: "February 10, 2016"
---
#Introduction

We want to analyze the characteristics of EDAV students in terms of 3 overall dimensions: 

  - Knowledge of statististical and computational tools
  - Confidence in analytic skills
  - Preference of text editor
  
The data we will be working with for this assignment consists of survey responses from 114 grad students, which are enrolled in one of the following programs:

  - IDSE
  - Data Science Certificate
  - Statistics
  - Others

##Cleaning the data
In order to work with clean and consistent data, we created a function called 'tinydata' that takes as input the csv file containing the raw responses, and returns a consistent data frame used across all the analyses performed.

Tinydata is stored as a script file named "tidydata.R" in our working directory that we need to source and run before we start the analysis.

In a nutshell, tidydata does the following:
 
  1.	Removes the columns that have no observations in them.
  2.	Regroups the columns in a logical manner.
  3.	Organizes the factor levels in the following columns
    a.	Waiting List Info
    b.	Gender
    c.	Program
    d.	Text Editors
  4.	Splits the "Experience with Tools"" column to multiple columns with a 1/0 value depending on if the skill was reported or not. Splitting the variable makes it tidy.

The distribution of respondents by program and gender is as follows:

```{r}
## INCLUDE distribution of respondents by program and gender
```

We continue our analysis by exploring the different toolkits that students from different masters typically possess.

##Tools
Respondents come from very diverse backgrounds and, as such, have learned and been exposed to very different sets of tools. 

There is a list of 16 tools, and each student was asked which of these he/she was comfortable using. As we can see in the following plot, "IDSE" and "Others" tend to know more tools on average

## Team: Verify last statement once averages are calculated
```{r}
##INCLUDE code that returns density of tools known by program (with average)
```

The next question we sought to answer was whether there are correlations between different sets of tools. In order to visualize this, we created a heatmap showing the correlation matrix for the top 10 tools. 

```{r}
#Show top 10 tools used (it's inside the heatmap function)
#Show heat map
```

As we can see, SQL, Github and Terminal have the strongest correlations, meaning that people that tend to know one of these tools, on average will know how to use the other two with a higher probability than the average student.

  - This result was somewhat expected. In general people with a stronger CS background will tend to know these three programs.

Interesting insights can also be observed by looking at the negative correlations seen between Python-R and between Matlab- Dropbox and Matlab-Google Drive. 

  - We are not surprised to see a negative correlation between R and Python. In general, they can be regarded as "substitute" tools, so people can work with one without ever having to learn the other.
  - The negative correlations between Matlab and Dropbox/Google drive are harder to explain. One potential explanation is that people that didn't know Matlab felt the need to fill in more tools, and so inputted dropbox or google drive.


#Skillsets

The next section of our work is dedicated to understand the skillsets that were reported by respondents.

The survey asked respondents to grade their level of confidence in 6 different dimensions which we call "Skills". The possible responses were "None", "A little", "Confident", and "Expert". As mentioned before, we assigned values from 0 to 3 to these levels to be able to aggregate data and compute averages. It is important to mention that this score is not perfect for several reasons, starting with the fact that these are self reported levels of confidence, and equal weight is being assigned to each skill.

The first exploratory analysis we did was to compute aggregated levels of confidence (called scores from here onward). The distribution of scores by skill looks as follows:

```{r}
#Insert pie chart here
```

Not surprisingly, we see that multivariate analysis has the highest overall confidence level. Now, we wish to have a bit more insight about how these confidence levels look when grouping by masters. 

```{r}
#Histograms showing skill scores distribution
```

##VERIFY next statement when data about average is available

Interestingly, Data Science Certificates and Statistics masters students seem to have the lowest level of confidence in their skills.

In order to understand which masters are confident with which skills, we took the analysis one step further to show with a plot this level of detail:

```{r}
#INSERT confidence lines by skills and masters here

##Bubble diagrams

```

##ADD A SUMMARY OF FINDINGS


#Text editor preference (NOT SURE IF WE WANT TO GO INTO THIS )



